---
title: 'Peer Graded Assignment: R markdown and Leaflet'
author: "RLC Velhorst"
date: "12-5-2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(dplyr)
set.seed(101)
```

# Developing Data Products Assignment Week 2: Tide Stations

This document is the report of the Peer Assessment project of week 2 from Coursera’s course Developing Data Products, as part of the Specialization in Data Science by John Hopkins University. It was built in RStudio, using its knitr functions. 

## Project Goal

The goal of your project is to demonstrate the student's capabilities of using R markdown and Leaflet. Using a R makrdown file, a HTML-file is built. The package Leaflet is used to create an interactive map in this HTML-file.  

The specific data in this exercise shows the location of worldwide tide stations.

## Data: Tide Station

The data used for this assignment originates from the Permanent Service for Mean Sea Level (PSMSL).
PSMSL is the global data bank for long term sea level change information from tide gauges and bottom pressure recorders.

This doument shows the location of all stations which are stored in de publicly available data: tide gauge data [Holgate et al., 2013; PSMSL, 2020].

```{r}

# data directory
datadir <- "./data"
if(!file.exists(datadir)){dir.create(datadir)}

# download zipped data
zipf = file.path(datadir, "rlr_annual.zip")
if(!file.exists(zipf)){
  fileUrl <- "https://www.psmsl.org/data/obtaining/year_end/2019/rlr_annual.zip"
  download.file(fileUrl,destfile = zipf)
  unzip(zipf,exdir="./data")
}

# file containing tide stations locations
fn_data <- "rlr_annual/filelist.txt"
dataf <-file.path(datadir, fn_data)

# station data
stations <- read.table(dataf, sep = ";",
                       col.names = c('id', 'lat', 'lon', 'name', 'coastline_code', 'station_code', 'quality'),
                       strip.white = TRUE
                       )

head(stations)
```

## Interactive map

```{r}

# Map stations,  Color by quality flag, 
# give station name on click,  start view around North Sea  and cluster stations


# Color by quality flag
getColor <- function(stations) {
  sapply(stations$quality, function(quality) {
  if(quality == 'N') {
    "green"
  } else {
    "red"
  } })
}

# Create icons using color
icons <- awesomeIcons(
  icon = 'none',
  markerColor = getColor(stations)
)


# Map stations
stations %>% 
  leaflet() %>% 
  setView(2, 54, zoom = 5) %>% # start view around North Sea
  addTiles() %>%
  addAwesomeMarkers(~lon, ~lat, icon=icons,
                    popup =stations$name, # give station name on click
                    clusterOptions = markerClusterOptions(maxClusterRadius = 25))



```

## Date: 12-5-2020


